#Patterns of Personal Movement  
##Week 2 Assignment
##Course: Reproducible Research
###By: Mindaugas Mozuraitis


###General activity patterns
```{r}
temp <- tempfile()
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)
week2data <- read.csv(unz(temp, "activity.csv"))
unlink(temp)
```

Histogram of the number of the number of steps taken per day:
```{r}
total_steps_per_day = aggregate(steps~date, data=week2data, sum)
hist(total_steps_per_day$steps)
```

The mean number of steps:
```{r}
round(mean(total_steps_per_day$steps))
```
The median number of steps:
```{r}
round(median(total_steps_per_day$steps))
```

The average daily activity pattern:
```{r}
mean_steps_per_int = aggregate(steps~interval, data=week2data, mean)
plot(steps~interval, data=mean_steps_per_int, type = "l")
```

The interval with the most steps on average:
```{r}
subset(mean_steps_per_int, steps==max(mean_steps_per_int$steps), select="interval")
```

The number of missing values in the dataset:
```{r}
sum(is.na(week2data$steps))
```

Missing values are susbtituted with the mean for that interval based on the days where that interval data is not missing:
```{r}
imputed = merge(week2data, mean_steps_per_int, by="interval") 
imputed$steps = ifelse(is.na(imputed$steps.x), imputed$steps.y, imputed$steps.x)  
```

The histogram of the total number of steps per day after the imputation:
```{r}
total_steps_per_day2 = aggregate(steps~date, data=imputed, sum)
hist(total_steps_per_day2$steps)
```

The mean number of steps after imputation: 
```{r}
round(mean(total_steps_per_day2$steps))
```
The median number of steps after imputation:
```{r}
round(median(total_steps_per_day2$steps))
```

###Activity patterns between weekdays and weekends
```{r}
imputed$dates = as.Date(imputed$date, format = "%Y-%m-%d")
imputed$weekdays = weekdays(imputed$dates)
imputed$weekpart = as.factor(ifelse(imputed$weekdays %in% c("Sunday", "Saturday"), "weekends", "weekdays"))
```

Activity during weekdays compared to the activity during weekends:
```{r}
mean_steps_per_wpint = aggregate(steps~interval+weekpart, data=imputed, mean)
library(ggplot2)
qplot(data=mean_steps_per_wpint, interval, steps, geom="line", facets=weekpart~.)
```